---
title: "Modeling"
author: "Elliot Wyman - 11317748"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(testit)
```

#### Import Data

```{r}
#read dataframes from csv
member.covariates <- read.csv("Data/member.covariates.csv")

attach(member.covariates)

head(member.covariates)
```

##### Models

1. modeling the response variable (e.g., make a donation) separately for each of the 2 groups (treatment and control)

###### split data

```{r}
tmp <- split(member.covariates, f = member.covariates$Treatment)

member.covariates.ctrl <- tmp[[1]]

member.covariates.treat <- tmp[[2]]

assert("Control group size is 900,000", nrow(member.covariates.ctrl) == 900000)

assert("Treatment group size is 100,000", nrow(member.covariates.treat) == 100000)
```

###### run logistic model predicting donation probability for control group (Pc)

```{r}
model_isdonate_control <- glm(donated ~ Treatment + Woman + Age + Salary + Education + City + Joined + read_rate + NineAmt + TenAmt + ElevenAmt + TwelveAmt + ThirteenAmt + FourteenAmt + FifteenAmt + SixteenAmt + SeventeenAmt + EighteenAmt + NineteenAmt + PC1 + PC2 + PC3 + PC4 + PC5 + Shares + Likes + Supporter, data=member.covariates.ctrl, family=binomial(link="logit"))

summary(model_isdonate_control)
```

###### run logistic model predicting donation probability for treatment group (Pt)

```{r}
model_isdonate_treatment <- glm(donated ~ Treatment + Woman + Age + Salary + Education + City + Joined + read_rate + NineAmt + TenAmt + ElevenAmt + TwelveAmt + ThirteenAmt + FourteenAmt + FifteenAmt + SixteenAmt + SeventeenAmt + EighteenAmt + NineteenAmt + PC1 + PC2 + PC3 + PC4 + PC5 + Shares + Likes + Supporter, data=member.covariates.treat, family=binomial(link="logit"))

summary(model_isdonate_treatment)
```

###### Calculate persuadable rate (Tau_c)

```{r}
prob_treat <- predict(model_isdonate_treatment, member.covariates)

prob_ctrl <- predict(model_isdonate_control, member.covariates)

tau_c = prob_treat - prob_ctrl
```

```{r}
results <- data.frame(member.covariates$ID, tau_c)

results$class[results$tau_c > 0] <- "Persuadable"
results$class[results$tau_c == 0] <- "Sure thing"
results$class[results$tau_c < 0] <- "Sleeping dog"

members.pers_sure <- split(member.covariates, results$tau_c >= 0)[[2]]

print(nrow(members.pers_sure))

head(members.pers_sure)
```

2. predict, according to the characteristics of the individual, the average amount donation amount (using only those who have made a donation)

###### split data

```{r}
members.donated <- split(member.covariates, member.covariates$donated)$`1`

head(members.donated)
```

```{r}
persuadable_threshold = 0

member.covariates$tau_c <- tau_c

members.predictions <- split(member.covariates, member.covariates$tau_c > persuadable_threshold)

members.persuadables <- members.predictions[[2]]

print(nrow(members.persuadables))

head(members.persuadables)
```

###### run linear regression model predicting donation amount

```{r}
model_donationamt_donators <- lm(TwentyAmt ~ Woman + Age + Salary + Education + City + Joined + read_rate + NineAmt + TenAmt + ElevenAmt + TwelveAmt + ThirteenAmt + FourteenAmt + FifteenAmt + SixteenAmt + SeventeenAmt + EighteenAmt + NineteenAmt + PC1 + PC2 + PC3 + PC4 + PC5 + Shares + Likes + Supporter, data=members.donated)

summary(model_donationamt_donators)
```

#### Results

##### predict donations 

```{r}
members.persuadables$pred_donation <- predict(model_donationamt_donators, members.persuadables)

members.pers_sure$pred_donation <- predict(model_donationamt_donators, members.pers_sure)
```

##### check if donators exist with estimated donation amount between $5 and $25

```{r}
sum(members.persuadables$pred_donation > 5) - sum(members.persuadables$pred_donation > 25) 
```
##### grab IDs of potential donators

```{r}
ids.dontcall <- subset(members.predictions[[1]], select=c(ID))

ids.call <- subset(members.predictions[[2]], select=c(ID))

results_collection <- rbind(data.frame(ids.dontcall, CALL = 0), data.frame(ids.call, CALL = 1))

head(results_collection)
```

##### Calculate net donations and prepare results data frame (donations of classes "Persuadables and Sure things") 

```{r}
sum(members.pers_sure$pred_donation)

results <- merge(results, members.pers_sure, by.x = "member.covariates.ID", by.y = "ID", all.x = TRUE)
```


```{r}
# write_csv() from the readr package is faster and never writes row names
library(readr)

#export dataframes to csv
write_csv(results_collection, file = "Data/Result_collection.csv")

#export dataframes to csv
write_csv(results, file = "Data/Results_summary.csv")
```

